---
title: "Data management report: Heart Failure Risk Across Cardio–Kidney–Metabolic Domains in Type 2 Diabetes"
date: "`r Sys.Date()`"
cover: static/ki_logo_vert_rgb.png
editor: source
format:
  pdf:
    documentclass: scrartcl
    template-partials:
      - "static/before-body.tex"
      - "static/_titlepage.tex"
    include-in-header: 
      - "static/in-header.tex"
    toc: true
    toc-depth: 3
    toc-title: Contents
    lof: false
    lot: true
    cap-location: top
    number-sections: true
    colorlinks: false
    keep-tex: false
bibliography: static/references.bib
csl: static/bmj.csl
link-citations: true
link-bibliography: true
nocite: '@*'
knitr:
  opts_chunk: 
    dev: cairo_pdf
    collapse: true
    comment: "" 
    echo: FALSE
    include: TRUE
    warning: FALSE
    message: FALSE
    fig.pos: "H"
    fig.path: "../output/figs/"
    fig.height: 6
    fig.width: 10
    R.options:
      knitr.graphics.auto_pdf: true
---

```{r}
#| label: set-up-load-data
options(knitr.kable.NA = "")

# load packages, globals and project specific functions
source(here::here("setup/setup.R"))

# load data
load(here("data/clean-data/ndrdata.RData"))
```           

\newpage

# Data

## Data sources

The data was linked using the Swedish personal identity number (PIN)[@pin]: 

- National Diabetes Register
- The Swedish Heart Failure Registry (SwedeHF) is an ongoing nationwide quality register started in 2000 
that includes in- and out-hospital patients with HF, regardless of EF and previous history of HF[@swedehf]. SHFDB version 4.2.3 was used for this study. 
- The National Prescribed Drug Register (The National Board of Health and Welfare) for information on medications[@npdr]
- The National Patient Register (The National Board of Health and Welfare) for additional comorbidities and outcomes[@npr]
- The Cause of Death Register[@dors] (The National Board of Health and Welfare) for outcomes death
- Longitudinal integrated database for health insurance and labour market studies (LISA)[@lisa] 
and the Register of the Total Population (Statistics Sweden) for socio-economic factors

## Ethics and Informed consent

The study was approved by the Swedish Ethical Review Authority, dnr 2021-04326. 

## Inclusion/exclusion criteria

```{r}
#| label: tbl-flow
#| tbl-cap: Information for flowchart
#| tbl-pos: "H"

default_kable(flow) %>%
  column_spec(1, width = "12cm")
```

First patient in: `r min(ndr$indexdtm)` and last patient in: `r max(ndr$indexdtm)`. 

Baseline is date of registered visit in the NDR. Note that the DM diagnosis can have occurred prior to this (see debutar). 

Follow-up is until emigration from Sweden, death or 2023-12-31, whichever comes first. 

## Definitions

{{< include src/vars.qmd >}}

## Overview

{{< include src/basetab.qmd >}}

# Reproducibility

## R code

The R code for all data handling is found: 

https://github.com/KIHeartFailure/ndr-ckd-hf. 
No data or output is stored in the repository. 

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

# References

::: {#refs}
:::
